---
title: "Module 7 ICE"
author: "Hayden Smedley"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
library(car)
library(dplyr)
library(ggplot2)

# Read file
brfss <- read.csv("BRFSS.csv")

```

## Data Recoding

The data is recoded for use in the various plots and tables to follow.

```{r}
# Recode Data
brfss$WEIGHT2[brfss$WEIGHT2 >= 7000] <- NA
brfss$DOCTDIAB[brfss$DOCTDIAB == 88] <- 0
brfss$DOCTDIAB[brfss$DOCTDIAB > 88] <- NA

brfss$MENTHLTH[brfss$MENTHLTH == 88] <- 0
brfss$MENTHLTH[brfss$MENTHLTH > 30] <- NA
brfss$EXERANY2[brfss$EXERANY2 == 1] <- "Yes"
brfss$EXERANY2[brfss$EXERANY2 == 2] <- "No"
brfss$EXERANY2[brfss$EXERANY2 == 7] <- "Don't Know"
brfss$EXERANY2[brfss$EXERANY2 == 9] <- NA
```

## Plot 1

This is a scatterplot that shows the positive correlation between weight and the number of health visits for diabetes in the last twelve months.

```{r}
plot(brfss$WEIGHT2, brfss$DOCTDIAB, 
     main = "Weight versus number of health visits for diabetes in the last year",
     xlab = "Weight in lbs", ylab = "Number of health visits for diabetes in the last year")
abline(lm(brfss$DOCTDIAB ~ brfss$WEIGHT2))

```

## Plot 2

This is a boxplot that shows the central tendancy of the number of mental health days in the last 30 days and if the respondent exercises outside of work.

```{r}
boxplot(MENTHLTH ~ EXERANY2, data = brfss,
        main = "Number of poor mental health days in the last 30 days by exercise",
        xlab = "Exercise outside of work", ylab = "Number of poor mental health days")

```

## Plot 3

This is a barplot that shows the counts of each category of exercise outside of work.

```{r}
barplot(table(brfss$EXERANY2),
        main = "Respondents by if exercises outside of work",
        xlab = "Exercises outisde of work", ylab = "Number of respondents")
```

## Plot 4

This is a table that shows the counts and frequency of each category of exercise outside of work.

```{r}
# Create data frame of EXERANY2 frequency and count tables
EXERANY2_df = data.frame(prop.table(table(brfss$EXERANY2)),table(brfss$EXERANY2))
EXERANY2_df = subset(EXERANY2_df, select = -Var1.1)
colnames(EXERANY2_df) <- c("Response", "Frequency", "Count")

# Print data frame
EXERANY2_df
```

## Plot 5

This is a scatterplot that shows the relationship between weight in lbs and the number of poor mental health days in the last 30 days colored by exercise outside of work.

```{r}
# Create data frame of WEIGHT2 and MENTHLTH grouped by EXERANY2
df = data.frame(brfss$WEIGHT2, brfss$MENTHLTH, brfss$EXERANY2)

# Plot data
plot_colors = c("black", "blue", "red")
plot(df$brfss.WEIGHT2, df$brfss.MENTHLTH, col = plot_colors[df$brfss.EXERANY2], pch = 19,
     main = "Number of poor mental health days by Weight in lbs colored by Exercise",
     xlab = "Weight in lbs", ylab = "Number of poor mental health days")
legend('topright',legend=levels(df$brfss.EXERANY2),col=plot_colors,pch=19,title='Exercise')
```

## Plot 6

This is a histogram that shows the frequncy of weight grouped by if the respondent exercises outside of work.

```{r}
# Create data frame of WEIGHT2 and EXERANY2
WEIGHT2_df = data.frame(brfss$WEIGHT2, brfss$EXERANY2)

# Plot data
ggplot(WEIGHT2_df, aes(x = brfss.WEIGHT2,group=brfss.EXERANY2,fill=brfss.EXERANY2)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 30) +
  labs(title = "Frequency of weight grouped by if exercises outside of work", fill = "Exercise") +
  xlab("Weight in lbs") + ylab("Frequency of weight")
```