---
title: "Logistic Regression - Predicting Diabetes Diagnosis"
author: ""
subtitle: "4/26/2020"
output: html_document
---
## Data Cleaning/Preparation
The first thing that I did was read in the diabetes csv. After that I created the height.si variable by multiplying the height variable by 0.0254 and the weight.si variable by multiplying the weight variable by 0.453592. Next I created the bmi_categorised variable by calculating the bmi for each person and then using that to sort them into the categories of underweight, normal, overweight, and obese with the rest as NA. Lastly I recoded the insurance and smoking variables to reflect their value labels.

```{r warning=FALSE, message=FALSE, results="hide"}
# Import data
setwd("C:/Users/hayde/Documents/GitHub/INST408F")
diabetes <- read.csv("Homework 4/diabetes.csv")

# Create height.si
diabetes$height.si <- diabetes$height*0.0254

# Create weight.si
diabetes$weight.si <- diabetes$weight*0.453592

# Create bmi_categorised
diabetes$bmi <- diabetes$weight.si/diabetes$height.si^2
diabetes$bmi_categorised <- cut(diabetes$bmi, 
                                breaks = c(0, 18.5, 25, 30, 100),
                                labels = c("underweight", "normal", "overweight", "obese"),
                                right = FALSE)

# Recode insurance variable
diabetes$insurance[diabetes$insurance == 0] <- "none"
diabetes$insurance[diabetes$insurance == 1] <- "government"
diabetes$insurance[diabetes$insurance == 2] <- "private"

# Recode smoking variable
diabetes$smoking[diabetes$smoking == 1] <- "current"
diabetes$smoking[diabetes$smoking == 2] <- "never"
diabetes$smoking[diabetes$smoking == 3] <- "ex-smoker"

```


## Exploratory Data Analysis
In the dataset there are 403 observations. Of those in the dataset 2.27% were underweight, 28.46% were of normal weight, 30.98% were overweight and 38.29% were obese by BMI. Additionally, 58.06% were Female and 41.94% were Male. 84.62% of those in the dataset did not have diabetes while 15.38% did. Lastly, the average age was 46.85 years old and the average BMI was 28.79.

```{r warning=FALSE, message=FALSE, results="hide"}
# Number of observations
nrow(diabetes)

# Precentages of BMI categories
prop.table(table(diabetes$bmi_categorised))*100

# Precentages of gender categories
prop.table(table(diabetes$gender))*100


# Precentages of diabetes categories
prop.table(table(diabetes$dm))*100

# Average age
mean(diabetes$age)

# Average BMI
mean(diabetes$bmi, na.rm = TRUE)
```


## Data Visualization
In Figure 1 the bar graph shows that the types of insurance are relatively even in their count but government insurance is the most common, followed by private and then no insurance least common. Figure 2 shows that people who have never smoked are by far the most common in the dataset, followed by current smokers and then ex-smokers. Figure 3 shows that there is a positive relationship between cholesterol and BMI, that is cholesterol tends to increase as BMI does. Lastly, Figure 4 shows the distribution of ages with around 40 being the most common but all ages between 25 and 65 are fairly similar in frequency.

```{r warning=FALSE, message=FALSE}
# Plot insurance
barplot(table(diabetes$insurance), 
        main = "Figure 1: Number of Types of Insurance", xlab = "Type of Insurance", ylab = "Count")

# Plot smoking
barplot(table(diabetes$smoking), 
        main = "Figure 2: Smoking Status", xlab = "Smoking Status", ylab = "Count")

# Plot BMI by Cholesterol
plot(diabetes$bmi, diabetes$chol, 
     main = "Figure 3: Comparison of Cholesterol by BMI", xlab = "Body Mass Index", ylab = "Cholesterol")

# Histogram of age
hist(diabetes$age, 
     breaks = 20,
     main = "Figure 4: Histogram of Age", xlab = "Age in Years", ylab = "Frequency")

```

## Logistic Regression Results
INSERT NARRATIVE HERE
```{r warning=FALSE, message=FALSE, results="hide"}
# insert logistic regression code


```

