---
title: "Logistic Regression - Predicting Diabetes Diagnosis"
author: ""
subtitle: "4/26/2020"
output: html_document
---
## Data Cleaning/Preparation
The first thing that I did was read in the diabetes csv. After that I created the height.si variable by multiplying the height variable by 0.0254 and the weight.si variable by multiplying the weight variable by 0.453592. Next I created the bmi_categorised variable by calculating the bmi for each person and then using that to sort them into the categories of underweight, normal, overweight, and obese with the rest as NA. Lastly I recoded the insurance and smoking variables to reflect their value labels.

```{r warning=FALSE, message=FALSE, results="hide"}
# insert import & data cleaning code here
setwd("C:/Users/hayde/Documents/GitHub/INST408F")
diabetes <- read.csv("Homework 4/diabetes.csv")

# Create height.si
diabetes$height.si <- diabetes$height*0.0254

# Create weight.si
diabetes$weight.si <- diabetes$weight*0.453592

# Create bmi_categorised
bmi <- diabetes$weight.si/diabetes$height.si^2
diabetes$bmi_categorised <- cut(bmi, 
                                breaks = c(0, 18.5, 25, 30, 100),
                                labels = c("underweight", "normal", "overweight", "obese"),
                                right = FALSE)

# Recode insurance variable
diabetes$insurance[diabetes$insurance == 0] <- "none"
diabetes$insurance[diabetes$insurance == 1] <- "government"
diabetes$insurance[diabetes$insurance == 2] <- "private"

# Recode smoking variable
diabetes$smoking[diabetes$smoking == 1] <- "current"
diabetes$smoking[diabetes$smoking == 2] <- "never"
diabetes$smoking[diabetes$smoking == 3] <- "ex-smoker"

```


## Exploratory Data Analysis
INSERT NARRATIVE HERE
```{r warning=FALSE, message=FALSE, results="hide"}
# insert exploratory data analysis code here



```


## Data Visualization
INSERT NARRATIVE HERE
```{r warning=FALSE, message=FALSE}
# insert data visualization code here



```

## Logistic Regression Results
INSERT NARRATIVE HERE
```{r warning=FALSE, message=FALSE, results="hide"}
# insert logistic regression code


```

